---
title: "DATA 2410 Group Project"
author: "Thu Nguyen and Minh Le"
date: "2025-04-10"
output: 
  html_document: 
    highlight: espresso
    theme: journal
---

# Charging Forward: A Data-Driven Look at Electric Vehicles Population (in Washington)
As the world transitions towards sustainable energy, electric vehicles (EVs) have emerged as a revolutionary force in the automotive industry. But which brands and models are actually driving this shift?

### Which are electric vehicles ?
Electric vehicles are referred to as battery electric vehicles (BEVs), plug-in hybrid electric vehicles (PHEVs), and hybrid electric vehicles (HEVs), which improves vehicle efficiency.

### Below is the data set provided by Washington State Department of Licensing:
```{r}
ev_data <- read.csv("Electric_Vehicle_Population_Data_1.csv", stringsAsFactors = FALSE)

library(DT)
datatable(ev_data[ , 1:10])
```

### Distribution of EV Types
```{r}
library(dplyr)
library(ggplot2)

ev_type_counts <- ev_data %>%
  count(Electric.Vehicle.Type)
ggplot(ev_type_counts, aes(x = "", y = n, fill = Electric.Vehicle.Type)) +
  geom_col(width = 1) +
  coord_polar("y") +
  labs(title = "Distribution of EV Types") +
  theme_void() +
  theme(legend.title = element_blank())
```

### Model Year vs. Electric Range
```{r}
ggplot(ev_data, aes(x = Model.Year, y = Electric.Range)) +
  geom_point(alpha = 0.5, color = "darkblue") +
  labs(title = "Model Year vs Electric Range",
       x = "Model Year",
       y = "Electric Range (miles)") +
  theme_minimal()
```

### What are the top 10 most common EV makes in Washington ?
```{r}
ev_data %>%
  count(Make, sort = TRUE) %>%
  top_n(10) %>%
  ggplot(aes(x = reorder(Make, n), y = n)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  labs(title = "Top 10 Most Common EV Makes",
       x = "Make", y = "Number of Vehicles")
```

### What are the most common EV models (top 10) in Washington ?
```{r}
ev_data %>%
  count(Model, sort = TRUE) %>%
  top_n(10) %>%
  ggplot(aes(x = reorder(Model, n), y = n)) +
  geom_col(fill = "forestgreen") +
  coord_flip() +
  labs(title = "Top 10 Most Common EV Models",
       x = "Model", y = "Count")
```
